Trending assessment for nowcasts, measurements, and forecasts

### 2024-03-07 sections/01-introduction.tex
Update on Overleaf.

The evaluation of trending is also interesting for the assessment of nowcasts.
In contrast to conventional forecasting methods, nowcasting methods focus on predictions for the present, the immediate future, and the recent past (\cite{banbura2013now}) and are now widely used in fields such as economics and medicine (\cite{bok2018macroeconomic, wolffram2023collaborative}).
Nowcasting has its origins in meteorology, and the methods were initially developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (\cite{browning1989nowcasting,schmid2019nowcasting}). In economics, nowcasting is used, for example, to predict statistics such as the current economic situation (for example, GDP), which are collected with low frequency and are available with a considerable time delay (\cite{banbura2013now}).
Nowcasting methods use indicators related to the target variable but can be collected at a higher frequency and published promptly to produce and update early estimates for the target variable during the relevant period (\cite{castle2017forecasting}). For example, in the context of forecasting GDP, we use economic indicators collected at a higher frequency from January to March 2024 to forecast GDP from the first quarter of 2024, before the official statistics are published.
In medicine, epidemic nowcasting is used during an ongoing epidemic to assess the current situation, taking into account the main pathogenic, epidemiological, clinical, and socio-behavioral factors (\cite{wu2021nowcasting}). 
The nowcasting method can be used, for example, to correct the daily Covid case numbers for events that have occurred but have not yet been reported before this data has been fully recorded at a central location (\cite{gunther2021nowcasting}). 
In this context, nowcasting is an estimation procedure in which the value of a target variable (e.g. from today) is estimated on the basis of current preliminary measurements, which is only finalised at a later point in time. 
At the time of nowcasting, estimates of the target variable can be made for the current time and for a past time; the predicted change in the target variable is then the difference between the two nowcasts. A nowcast is consistent with the evolution of the target variable if the difference between the actual values at the current time and the actual values at a past time has the same sign as the predicted change. Like forecasts, nowcasts are often evaluated and compared in the literature on the basis of performance measures such as the root mean absolute error (\cite{gunther2021nowcasting}) and the aspect of trending ability is not considered in the literature to the best of our knowledge. However, appropriate measures for assessing trending ability should be constructed, because a nowcast with strong trending ability provides crucial information about the development of the target variable (for example, the number of cases of an epidemic) and can be a solid basis for decisions on the introduction or cancellation of policy measures.

As outlined above, the ability to take into account the trendability of a method depending on the area of application, be it measurements, predictions and forecasts, is of high importance. 
However, regardless of the application domain, most conventional method comparisons and evaluations consider local absolute differences in different ways without directly assessing whether the correct direction of change is output.

>>>>

The evaluation of trending is also interesting for the assessment of nowcasts.
In contrast to conventional forecasting methods, nowcasting methods focus on predictions for the present, the immediate future, and the recent past (\cite{banbura2013now}) and are now widely used in fields such as economics and medicine (\cite{bok2018macroeconomic, wolffram2023collaborative}).
Nowcasting has its origins in meteorology, and the methods were initially developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (\cite{browning1989nowcasting,schmid2019nowcasting}). In economics, nowcasting is used, for example, to predict statistics such as the current economic situation (for example, GDP), which are collected with low frequency and are available with a considerable time delay (\cite{banbura2013now}).
Nowcasting methods use indicators that are related to the target variable and collected at a higher frequency.
Thus, the nowcasts can produce early estimates for the target variable during the relevant period (\cite{castle2017forecasting}). For example, in forecasting GDP, we use economic indicators collected at a higher frequency from January to March 2024 to forecast GDP from the first quarter of 2024 before the official statistics are published.
In medicine, epidemic nowcasting is used to assess the current situation during an ongoing epidemic, considering the main pathogenic, epidemiological, clinical, and socio-behavioral factors (\cite{wu2021nowcasting}). 
For example, the nowcasting method can correct the daily COVID case numbers for events that have occurred but have not yet been reported before this data has been fully recorded at a central location (\cite{gunther2021nowcasting}). 
In this context, nowcasting is an estimation procedure in which the value of a target variable (for example, from today) is estimated based on current preliminary measurements, which are only finalized later. 
At the time of nowcasting, estimates of the target variable can be made for the current time and for a past time; the predicted change in the target variable is then the difference between the two nowcasts. A nowcast is consistent with the evolution of the target variable if the difference between the actual values at the current time and the actual values at a past time has the same sign as the predicted change. Like forecasts, nowcasts are often evaluated and compared in the literature based on performance measures such as the root mean absolute error (\cite{gunther2021nowcasting}) or probabilistic measures \parencite{Wolffram2023}. The aspect of trending ability is not considered in the literature to the best of our knowledge. However, appropriate measures for assessing trending ability should be constructed because a nowcast with strong trending ability provides crucial information about the development of the target variable, for example, the number of cases of an epidemic, and can be the foundation of decisions on the introduction or cancellation of policy measures.

As outlined above, trending evaluation is crucial for measurements, predictions, and forecasts. 
However, regardless of the application domain, most conventional method comparisons and evaluations consider local absolute differences in different ways without directly assessing whether the correct direction of change is output.
